Forecasting biodiversity in breeding birds using best practices

### 2017-12-08 manuscript/reviewer_reply.Rmd

*Line 269: See my above comments about spatially correlated random effects / GMRF models. One way to confirm these other approaches are overkill would be to make some maps of the estimated random effects, and fit some simple models to estimate variogram parameters – e.g. even gls() in R can do this*

>See our comments about spatial autocorrelation above

*Line 276: Is one of the potential reasons that richness is stationary that presence/absence on the BBS varies less than other measures – maybe abundance, or other measures of diversity are more variable?*

>We agree that this is likely. We make this point on [[lines]]

@@

*Line 294: Is the narrow predictive distribution for the stacked SDM also why this exhibits the strongest trend in decreasing coverge as a function of forecast length (Fig 5)?*

>Yes, exactly. More and more observations fall outside the narrow prediction intervals.

@@

*Line 329 – maybe reword, you use forecast 3x in this sentence*

*Line 414: Agreed – doing this same kind of forecasting competition with compositional data, or other species diversity measures (Simpson’s) would be interesting*

>We are currently working on a few projects along these lines.

@@

*Line 461: Another thread with conservation implications is that back in the 1990s, it was pretty common for single species PVA models to be forecasted 100-200 years. This seems less common, in part because some predictions may have not performed well, and others are on a time horizon that can’t be evaluated. In Steve Beissinger’s PVA book he cites Goldwasser (2000) “Variability and measurement error in extinction risk analysis...” for this point (p 133).*

>We agree that this is an apt comparion. However, we are unfamiliar with this literature and UF does not have access to the paper you cited.

### 2017-12-08 manuscript/reviewer_reply.Rmd

*Line 414: Agreed – doing this same kind of forecasting competition with compositional data, or other species diversity measures (Simpson’s) would be interesting*

>We are currently working on a few projects along these lines.

*Line 461: Another thread with conservation implications is that back in the 1990s, it was pretty common for single species PVA models to be forecasted 100-200 years. This seems less common, in part because some predictions may have not performed well, and others are on a time horizon that can’t be evaluated. In Steve Beissinger’s PVA book he cites Goldwasser (2000) “Variability and measurement error in extinction risk analysis...” for this point (p 133).*

>We agree that this is an apt comparison. However, we are unfamiliar with this literature and UF does not have access to the paper you cited. The existing citations should prove sufficient to justify these points.

@@

#### Comments for the author

*In this paper, the authors use the incredibly rich data from the Breeding Bird Survey to examine how well various models perform at predicting bird richness at sites throughout North America. They train models on data from 1982-2003 and attempt to predict data from 2004-2013. They examine the performance of six different model types, two of which are null or baseline models.*

*I find it a bit surprising that the authors think that any of these richness models could predict a high level of accuracy for theses ten years. I think of richness estimations as a very coarse technique for looking at broad shifts in richness over relatively long time periods rather than these 5-species (~5-7%) shifts in richness on annual bases indicated in Figure 2. The authors don’t provide a lot of their raw data on sites other than as examples, but my impression is that they are pursuing an unrealistic level of accuracy and detail. That being said, given what the paper does, it’s okay. The models seem to be performed well. The authors give all of the keywords to indicate that they know what they are doing.*

>We agree that predicting smaller, transient shifts in richness may be more difficult than predicting large, sustained ones. Nevertheless, we believe that our results have implications for ecologists' longer-term predictions of declining richness, as discussed on lines [[XX]].

### 2017-12-08 manuscript/draft.Rmd

Biodiversity forecasts are important for conservation, management, and
evaluating how well current models characterize natural systems. While the
number of forecasts for biodiversity is increasing, there is little information
available on how well these forecasts work. Most biodiversity forecasts are not
evaluated to determine how well they predict future diversity, fail to account
for uncertainty, and do not use time-series data that captures the actual
dynamics being studied. We addressed these limitations by using best practices
to explore our ability to forecast the species richness of breeding birds in
North America. We used hindcasting to evaluate six different modeling approaches
for predicting richness. Hindcasts for each method were evaluated annually for a
decade at `r numbers$N_predict_sites` sites distributed throughout the
continental United States. While each model could explain most of the variance
in richness, none of them consistently outperformed a baseline model that
predicted constant richness at each site. In particular, we found no evidence
that current methods (such as species distribution models) can successfully turn
spatial data into useful temporal predictions about biodiversity at decadal
time-scales. The best practices implemented in this study directly influence the
forecasts, the relative performance of different modeling approaches, and the
conclusions about the current state of biodiversity forecasting . To facilitate
the rapid improvement of biodiversity forecasts, we emphasize the value of
specific best practices in making forecasts and evaluating forecasting methods.

>>>>

Biodiversity forecasts are important for conservation, management, and
evaluating how well current models characterize natural systems. While the
number of forecasts for biodiversity is increasing, there is little information
available on how well these forecasts work. Most biodiversity forecasts are not
evaluated to determine how well they predict future diversity, fail to account
for uncertainty, and do not use time-series data that captures the actual
dynamics being studied. We addressed these limitations by using best practices
to explore our ability to forecast the species richness of breeding birds in
North America. We used hindcasting to evaluate six different modeling approaches
for predicting richness. Hindcasts for each method were evaluated annually for a
decade at `r numbers$N_predict_sites` sites distributed throughout the
continental United States. All models explained more than 50% of the variance in
richness, but none of them consistently outperformed a baseline model that
predicted constant richness at each site. The best practices implemented in this
study directly influenced the forecasts and evaluations. Stacked species
distribution models and "naive" forecasts produced poor estimates of uncertainty
and accounting for this resulted in these models dropping in the relative
performance compared to other models. Accounting for observer effects improved
model performance overall, but also changed the rank ordering of models because
it did not improve the accuracy of the "naive" model. Considering the forecast
horizon revealed that the prediction accuracy decreased across all models as the
time horizon of the forecast increased. To facilitate the rapid improvement of
biodiversity forecasts, we emphasize the value of specific best practices in
making forecasts and evaluating forecasting methods.
